---
title: "Sidewalk Damage and Population Change"
author: "Ashley Culver"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, error=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load in data
data_2015 <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/data_2015.rds")
data_2005 <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/data_2005.rds")
data_1995 <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/data_1995.rds")
pop_data <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/population.rds")
```

```{r}
#load in packages
library(tidyverse)
library(ggplot2)
library(scales)

#create tidy dataset
slim_sidewalk <- data_2015 %>% 
  select(sidewalk, borough)
```
Sidewalk condition 2015
1) damage by borough
- damage broken down proportionally by borough
2) damage by neighborhood

Percentage of damaged sidewalk in each borough, 2015
```{r}
library(ggplot2)
# prep for plot
damage_sidewalk <- slim_sidewalk %>%
  group_by(sidewalk, borough) %>% 
  summarize(count = n())%>%
  filter(sidewalk == "Damage")%>%
  mutate(percent_damage = count/sum(count))

damage_sidewalk <-damage_sidewalk%>% 
  arrange(desc(borough)) %>%
  mutate(ypos = cumsum(percent_damage)- 0.5*percent_damage )
```

Plot: How is all damaged sidewalk in NYC proportioned across boroughs?
```{r}
#damage plot (How is all damaged sidewalk in NYC proportioned across boroughs?)
p<- ggplot(damage_sidewalk, aes(x="", y = percent_damage, fill=borough))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y")+
  theme_void()+
  geom_text(aes(y = ypos, label = percent(percent_damage)), color = "white")+
  labs(fill = "Borough")+
  ggtitle("Proportion of NYC Sidewalk Damage per Borough")

p
```

Sidewalk Damage by Borough
```{r}
data_2015_damage <- slim_sidewalk %>% 
  group_by(borough, sidewalk) %>% 
  summarize(count = n()) %>%
  mutate(percent_damage = count/sum(count)) %>%
  filter(sidewalk=="Damage") %>%
  group_by(borough)
```

Plot: What percentage of sidewalk damaged in each borough?
```{r}
#damage plot (what percentage of sidewalk damaged in each borough?)
ggplot(data_2015_damage, aes(reorder(borough, percent_damage),100*percent_damage)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  scale_y_continuous(name="% Sidewalk Damaged") +
  coord_flip() +
  labs(x = "Borough", fill = "Borough") + 
  ggtitle("Sidewalk Damage by Borough, 2015")
```
Change in population by neighborhood 2000 - 2010
```{r}
#top 10 neighborhoods by population total
pop_data_2000 <- pop_data %>%
  filter(year == 2000) %>%
  arrange(desc(population)) %>%
  top_n(10, population)

pop_data_2010 <- pop_data %>%
  filter(year == 2010) %>%
  arrange(desc(population)) %>%
  top_n(10, population)

```

Plot: top 10 neighborhoods by population total, 2010
```{r}
#plot of top 10 neighborhoods in NYC by largest population
ggplot(pop_data_2010, aes(reorder(nta_name, population),population)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  scale_y_continuous(name="Population", labels = scales::comma) +
  coord_flip() +
  labs(x = "Neighborhood", fill = "Borough") + 
  ggtitle("Top 10 NYC Neighborhoods by Largest Population, 2010")
```

top 5 neighborhoods by population per borough
```{r}
pop_data_2000 <- pop_data %>%
  group_by(borough) %>%
  filter(year == 2000) %>%
  arrange(desc(population), .by_group = TRUE) %>%
  top_n(5, population)

pop_data_2010 <- pop_data %>%
  group_by(borough) %>%
  filter(year == 2010) %>%
  arrange(desc(population), .by_group = TRUE) %>%
  top_n(5, population)
```

Plot: top 5 neighborhoods by population by borough, 2010
```{r}
#plot of top 5 neighborhoods per borough by population
ggplot(pop_data_2010, aes(reorder(nta_name, population),population)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  scale_y_continuous(name="Population", labels = scales::comma) +
  coord_flip() + 
  labs(x = "Neighborhood", fill = "Borough") + 
  ggtitle("Top 5 Neighborhoods/Borough by Population, 2010")
```

top 10 neighborhoods by population growth by borough
```{r}
#add population change and growth rate to data
#remove parks/cemeteries, jails, and airports
pop_decade <- pop_data %>%
  group_by(nta_code) %>%
  filter(!nta_code%in% c("BX99", "BX98", "BK99", "BK98", "MN99", "MN98", "QN99", "QN98", "SI99", "SI98")) %>%
  mutate(pop_change = population[year==2010] - population[year==2000]) %>%
  mutate(pop_growth = pop_change/population[year==2000]) 
   
#look at growth rate by borough
pop_growth_borough <- pop_decade %>%
  group_by(borough) %>%
  summarize(pop_change_borough = sum(pop_change)/2/sum(population[year==2000]))

pop_growth_borough_nta <- pop_decade %>%
  group_by(borough) %>%
  filter(year == 2010) %>%
  arrange(desc(pop_growth), .by_group = TRUE) %>%
  top_n(5, pop_growth)

```

Plot: Population growth by borough, 2000 - 2010
```{r}
ggplot(pop_growth_borough, aes(reorder(borough, pop_change_borough), 100*pop_change_borough)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  #scale_y_continuous(name="Population", labels = scales::comma) +
  coord_flip() + 
  labs(x = "Neighborhood", y = "% Population Growth 2000 - 2010", fill = "Borough") + 
  ggtitle("Population Growth by Borough, 2000 - 2010")
```

Plot: Population growth by neighborhood (top 5 per borough), 2000 - 2010
```{r}
ggplot(pop_growth_borough_nta, aes(reorder(nta_name, pop_growth), 100*pop_growth)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  coord_flip() + 
  labs(x = "Neighborhood", y = "% Population Growth 2000 - 2010", fill = "Borough") + 
  ggtitle("Top 5 Neighborhoods per Borough by Population Growth, 2000 - 2010")
```

top 10 neighborhoods by population growth total
```{r}
#look at growth rate by neighborhood
pop_growth_neighborhood <- pop_decade %>%
  filter(year==2010) %>%
  arrange(desc(pop_growth)) %>%
  top_n(10, pop_growth)

pop_growth_neighborhood_f <- pop_growth_neighborhood[1:10,]
```

Plot: Population growth by neighborhood (all NYC), 2000 - 2010
```{r}
ggplot(pop_growth_neighborhood_f, aes(reorder(nta_name, pop_growth), 100*pop_growth)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  coord_flip() + 
  labs(x = "Neighborhood", y = "% Population Growth 2000 - 2010", fill = "Borough") + 
  ggtitle("Top 10 NYC Neighborhoods by Population Growth, 2000 - 2010")
```


