---
title: "Sidewalk Damage and Population Change"
author: "Ashley Culver"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, error=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load in data
data_2015 <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/data_2015.rds")
data_2005 <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/data_2005.rds")
data_1995 <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/data_1995.rds")
pop_data <- readRDS("~/Documents/GitHub/fall2019-proj2--sec1-grp2/data/population.rds")

```

```{r}
#load in packages
library(tidyverse)
library(ggplot2)
library(scales)

#create tidy dataset
slim_sidewalk <- data_2015 %>% 
  select(tree_id, sidewalk, borough, block_id, tree_dbh, curb_loc, status, health, spc_common, steward, user_type, problems, postcode, nta, nta_name, latitude, longitude)
```
Sidewalk condition 2015
1) damage by borough
- damage broken down proportionally by borough
2) damage by neighborhood
3) damage by tree type
4) damage by tree diameter

Sidewalk condition across years. 

1) Problems persist for same trees in which areas?
2) STILL NEED TO ADD CHART NAME TO BELOW FIGURE
```{r}
#look at sidewalk data by borough
data_2015 %>% 
  group_by(borough, sidewalk) %>% 
  summarize(count = n())%>%
  mutate(percent = 100*count/sum(count)) #percent trees per borough by sidewalk status

data_2015 %>% 
  group_by(sidewalk, borough) %>% 
  summarize(count = n())%>%
  mutate(percent = 100*count/sum(count)) #percent trees per sidewalk status by borough
```
Percentage of damaged sidewalk in each borough, 2015
```{r}
library(ggplot2)
# plot
damage_sidewalk <- slim_sidewalk %>%
  select(sidewalk, borough) %>% 
  group_by(sidewalk, borough) %>% 
  summarize(count = n())%>%
  filter(sidewalk == "Damage")%>%
  mutate(percent_damage = count/sum(count))

damage_sidewalk <-damage_sidewalk%>% 
  arrange(desc(borough)) %>%
  mutate(ypos = cumsum(percent_damage)- 0.5*percent_damage )

p<- ggplot(damage_sidewalk, aes(x="", y = percent_damage, fill=borough))+
  geom_bar(width = 1, stat = "identity")+
  coord_polar("y")+
  theme_void()+
  geom_text(aes(y = ypos, label = percent(percent_damage)), color = "white")+
  labs(fill = "Borough")+
  ggtitle("Proportion of NYC Sidewalk Damage per Borough")

p
```
Damaged sidewalk by tree type
```{r}
#start with problem areas in each dataset
damage_sidewalk_tree <- data_2015 %>%
  select(tree_dbh, sidewalk, spc_common, block_id, borough, nta_name, latitude, longitude) %>%
  filter(sidewalk == "Damage")
  

damage_sidewalk_tree_type  <- damage_sidewalk_tree %>%
  group_by(spc_common) %>%
  summarize(count = n())

damage_sidewalk_tree_borough <- damage_sidewalk_tree %>%
  group_by(borough) %>%
  summarize(count = n())

damage_sidewalk_tree_nta_name <- damage_sidewalk_tree %>%
  group_by(nta_name) %>%
  summarize(count = n())

damage_sidewalk_tree_dbh <- damage_sidewalk_tree %>%
  group_by(tree_dbh) #%>%
 # summarize(count = n(),
            
```

Sidewalk Damage by Borough
```{r}
data_2015_damage <- data_2015 %>% 
  group_by(borough, sidewalk) %>% 
  summarize(count = n()) %>%
  mutate(percent_damage = count/sum(count)) %>%
  filter(sidewalk=="Damage") %>%
  group_by(borough)

ggplot(data_2015_damage, aes(reorder(borough, percent_damage),100*percent_damage)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  scale_y_continuous(name="% Sidewalk Damaged") +
  coord_flip() +
  labs(x = "Borough", fill = "Borough") + 
  ggtitle("Sidewalk Damage by Borough, 2015")
```
Change in population by neighborhood 2000 - 2010
```{r}
#top 10 neighborhoods by population total
pop_data_2000 <- pop_data %>%
  filter(year == 2000) %>%
  arrange(desc(population)) %>%
  top_n(10, population)

pop_data_2010 <- pop_data %>%
  filter(year == 2010) %>%
  arrange(desc(population)) %>%
  top_n(10, population)

```
top 10 neighborhoods by population total, 2010
```{r}
ggplot(pop_data_2010, aes(reorder(nta_name, population),population)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  scale_y_continuous(name="Population", labels = scales::comma) +
  coord_flip() +
  labs(x = "Neighborhood", fill = "Borough") + 
  ggtitle("Top 10 NYC Neighborhoods by Largest Population, 2010")
```
top 5 neighborhoods by population per borough
```{r}
pop_data_2000 <- pop_data %>%
  group_by(borough) %>%
  filter(year == 2000) %>%
  arrange(desc(population), .by_group = TRUE) %>%
  top_n(5, population)

pop_data_2010 <- pop_data %>%
  group_by(borough) %>%
  filter(year == 2010) %>%
  arrange(desc(population), .by_group = TRUE) %>%
  top_n(5, population)
```

top 5 neighborhoods by population by borough, 2010
```{r}
ggplot(pop_data_2010, aes(reorder(nta_name, population),population)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  scale_y_continuous(name="Population", labels = scales::comma) +
  coord_flip() + 
  labs(x = "Neighborhood", fill = "Borough") + 
  ggtitle("Top 5 Neighborhoods per Borough by Largest Population, 2010")
```

top 10 neighborhoods by population growth by borough
```{r}
#add population change and growth rate to data
#remove parks/cemeteries, jails, and airports
pop_decade <- pop_data %>%
  group_by(nta_code) %>%
  filter(!nta_code%in% c("BX99", "BX98", "BK99", "BK98", "MN99", "MN98", "QN99", "QN98", "SI99", "SI98")) %>%
  mutate(pop_change = population[year==2010] - population[year==2000]) %>%
  mutate(pop_growth = pop_change/population[year==2000]) 
   

#look at growth rate by borough
pop_growth_borough <- pop_decade %>%
  group_by(borough) %>%
  summarize(pop_change_borough = sum(pop_change)/2/sum(population[year==2000]))

pop_growth_borough_nta <- pop_decade %>%
  group_by(borough) %>%
  filter(year == 2010) %>%
  arrange(desc(pop_growth), .by_group = TRUE) %>%
  top_n(5, pop_growth)

```

```{r}
ggplot(pop_growth_borough, aes(reorder(borough, pop_change_borough), 100*pop_change_borough)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  #scale_y_continuous(name="Population", labels = scales::comma) +
  coord_flip() + 
  labs(x = "Neighborhood", y = "% Population Growth 2000 - 2010", fill = "Borough") + 
  ggtitle("Population Growth by Borough, 2000 - 2010")
```

```{r}
ggplot(pop_growth_borough_nta, aes(reorder(nta_name, pop_growth), 100*pop_growth)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  coord_flip() + 
  labs(x = "Neighborhood", y = "% Population Growth 2000 - 2010", fill = "Borough") + 
  ggtitle("Top 5 Neighborhoods per Borough by Population Growth, 2000 - 2010")
```

top 10 neighborhoods by population growth total
```{r}
#look at growth rate by neighborhood
pop_growth_neighborhood <- pop_decade %>%
  filter(year==2010) %>%
  arrange(desc(pop_growth)) %>%
  top_n(10, pop_growth)

pop_growth_neighborhood_f <- pop_growth_neighborhood[1:10,]
```

```{r}
ggplot(pop_growth_neighborhood_f, aes(reorder(nta_name, pop_growth), 100*pop_growth)) +
  geom_bar(stat = "identity", aes(fill = borough)) +
  coord_flip() + 
  labs(x = "Neighborhood", y = "% Population Growth 2000 - 2010", fill = "Borough") + 
  ggtitle("Top 10 NYC Neighborhoods by Population Growth, 2000 - 2010")
```


